---
title: "celiotomy retrospective analysis 2023"
author: "Gregg Griffenhagen"
date: "2023-08-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(magrittr)
library(logistf)

data <- read.csv('celio_retro_LP_GG_23.csv', 
                 nrows=608,
                 colClasses=c("character","integer","factor",
                                                          "numeric", "numeric", "factor",
                                                          "factor", "factor", "factor",
                                                          "factor", "factor", "factor",
                                                          "factor", "factor", "factor",
                                                          "numeric", "integer", "factor",
                                                          "integer", "integer", "factor",
                                                          "factor", "factor", "factor"
                                                          ),
                 na.strings=c("", " "))
data$date_pres <- as.Date(data$date_pres, format="%m/%d/%y")
str(data)
```

#Model primary outcome of incisional infection vs variables that make sense:

```{r}
model_all <- logistf(incis_infect ~ enterot + bowel_resect + preop_antibio + intraop_antibio +
                       anes_time + recov_time + recov_qual, data = data)
summary(model_all)
```
##Note that none of the variables in the full model are significant. Backwards elimination removed all of the variables and resulted in a model of (incisional infection ~ 1).
## The next step would be to trial individual sets of variables that may be significant

### First try enterotomy or resection:
```{r}
model1 <- logistf(incis_infect ~ enterot + bowel_resect,
                  data = data)
summary(model1)
```
### Antibiotic use:
```{r}
model2 <- logistf(incis_infect ~ preop_antibio + intraop_antibio,
                  data = data)
summary(model2)
```
### Intraoperative antibiotic type or timing?
```{r}
model3 <- logistf(incis_infect ~ intraop_antibio_type + intraop_antibio_time,
                  data = data)
summary(model3)
```
### Anesthesia variables?
```{r}
model4 <- logistf(incis_infect ~ anes_time + recov_time + recov_qual,
                  data = data)
summary(model4)
```
### Association with NSAID use?
```{r}
model5 <- logistf(incis_infect ~ postop_nsaid_num + postop_nsaid_days,
                  data = data)
summary(model5)
```

## None of these produced significant independent predictors without possible confounding. The use of multiple days of NSAIDs is likely due to the infection, not the cause of the infection.

# How about modeling postop reflux?
```{r}
model6 <- logistf(postop_reflux ~ anes_time + bowel_resect + enterot + preop_antibio +
                    intraop_antibio + postop_antibio_days + postop_antibio_addnl +
                    postop_nsaid_num + postop_nsaid_days + postop_lido +
                    postop_alpha2 + postop_butor + postop_ket, data=data)
summary(model6)
```

